Calibrating seed-based alignment heuristics with Sesame

### 2020-03-09 manuscript.tex
Add section "Practical considerations"

We need to know whether a nucleotide is a sequencing error, but this time
we also need to know its phase in the repeated cycles of skipped
positions. For this, we define the skip-$n$ alphabet $\mathcal{A}_n =
\{\square, |, \Downarrow_0, \Downarrow_1, \ldots, \Downarrow_n \}$. Again,
the symbol $\square$ represents a correct nucleotide and the symbol $|$ is
a terminator added at the end of the read. The symbols $\Downarrow_j$ ($j
\in \mathbb{N}$) represent sequencing errors and $j$ indicates the number
of nucleotides till the next non-skipped position (\textit{i.e.} $j=0$ for
nucleotides immediately before a non-skipped position and $j=n$ for
nucleotides at a non-skipped position).

@@

Finally, the expression of $M_n(z)$ the transfer matrix of reads in the
skip-$n$ alphabet is
\begin{equation*}
\begin{blockarray}{cccccc}
   & \Dn{0} & \Dn{1} & \ldots & \Dn{n} & \nd \\
\begin{block}{c[ccccc]}
\Dn{0} & H_{0,0}(z) & H_{0,1}(z) & \ldots & H_{0,n}(z) & J_0(z) \\
\Dn{1} & H_{1,0}(z) & H_{1,1}(z) & \ldots & H_{1,n}(z) & J_1(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\Dn{n} & H_{n,0}(z) & H_{n,1}(z) & \ldots & H_{n,n}(z) & J_n(z) \\
\nd & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray}.
\end{equation*}

@@

By construction, the Taylor expansion of the top right term of the matrix
$M_n(z)\cdot(I-M_n(z))^{-1}$ contains the the probabilities of interest.
More specifically, $a_k$, the coefficient of $z^k$ in the series is the
probability that a read of size $k$ contains no skip-$n$ seed of minimum
size $\gamma$.

But $M_n(z)$ is too complex to find a closed expression for $M_n(z) \cdot
(I-M_n(z))^{-1}$. Instead, we return to the definition $M_n(z) + M_n(z)^2
+ \ldots$ and show in appendix~\ref{app_skip} that the coefficient of
interest, $a_k$, only depends on the first $k+1$ terms of the sum. So for
reads of size $k$ or lower, we only need to compute the matrix $M_n(z) +
M_n(z)^2 + \ldots + M_n(z)^{k+1}$ and work out the Taylor expansion of the
top right term.

@@

Using the formula above, we can thus bound the probability that a read has
$s+1$ segments or more. We compute $M_n(z) + M_n(z)^2 + \ldots
+M_n(z)^{s+1}$ where the weighted generating functions have been replaced
by truncated polynomials and we extract the top right entry. When the
bound is lower than a set fraction $\varepsilon$ of the current value of
$a_k$, we stop the computations. Typically $\varepsilon = 0.01$ so this
method ensures that the probabilities that a read of size $k$ has no
on-target skip seed are accurate to within 1\%.

@@

We now turn our attention to the problem of computing the probability that
the seeding process is off target when using exact seeds --- recall from
section~\ref{sec:twotypes} that off-target seeding means that the
candidate set contains a duplicate but not the target.

If there is $N = 0$ duplicate, seeding cannot be off target, it can only
be on target or null. So from here we assume that the target has $N \geq
1$ duplicates. Let $S_0$ denote the event that there is an on-target seed
and let $S_j$ denote the event that there is a seed for the $j$-th
duplicate. We are thus interested in computing $P\big(\overline{S}_0 \cap
(S_1 \cup \ldots \cup S_N)\big)$, where $\overline{S}_j$ denotes the
complement of the event $S_j$. For this, we first observe that
\begin{equation}
\label{eq:pcomp}
P\big(\overline{S}_0 \cap (S_1 \cup \ldots \cup S_N)\big) =
P(\overline{S}_0) - P\big(\overline{S}_0 \cap \overline{S}_1 \cap \ldots
\cap \overline{S}_N\big).
\end{equation}

Since the duplicates are assumed to evolve independently of each other and
through the same mutagenesis process, the events are conditionally
independent (we do not have full independence because $S_0$ only depends
on sequencing errors whereas the events $S_j$ depend on sequencing errors
and on the divergence between duplicates). We can thus write
\begin{eqnarray}
\notag
P(\overline{S}_0 \cap \ldots \cap \overline{S}_N) &=&
  P(\overline{S}_0) \cdot P(\overline{S}_1 \cap \ldots \cap
  \overline{S}_N | \overline{S}_0) \\
\label{eq:Pnull}
  &=& P(\overline{S}_0) \cdot P(\overline{S}_1 | \overline{S}_0)^N =
  P(\overline{S}_0) \cdot \left( \frac{P(\overline{S}_0 \cap
\overline{S}_1)}{P(\overline{S}_0)} \right)^N.
\end{eqnarray}
Combinging the two equations above, we obtain
\begin{equation}
\label{eq:Poff}
P\big(\overline{S}_0 \cap (S_1 \cup \ldots \cup S_N)\big) =
P(\overline{S}_0) - P(\overline{S}_0) \cdot \left( \frac{P(\overline{S}_0
\cap \overline{S}_1)}{P(\overline{S}_0)} \right)^N.
\end{equation}

Hence, the probability that seeding is off target is a function of two
quantitites. We have already seen how to compute $P(\overline{S}_0)$ in
section~\ref{sec:example_exact} using recursive expression
(\ref{eq:recur}). We now need to find a way to compute $P(\overline{S}_0
\cap \overline{S}_1)$.

@@

The event $\overline{S}_0 \cap \overline{S}_1$ is that the read contains
no seed for either the target or for the first duplicate --- numbering is
arbitrary here, we simply chose a duplicate and stick to it. As
illustrated in section~\ref{sec:symbolic}, we first recode the reads using
a specialized alphabet to simplify the problem.

It will be useful to consider a more general problem where we
have two sequences of interest labelled $+$ and $-$. The $+$ sequence
stands for the target and that the $-$ sequence stands for its duplicate.
We then define the dual alphabet $\tilde{\mathcal{A}}_0 = \{\square, |,
\downarrow_{/1}^-, \downarrow_{/2}^-, \ldots, \downarrow_{/1}^+,
\downarrow_{/2}^+, \ldots, \Downarrow\}$. The symbols $\downarrow_{/i}^-$
signify that the nucleotide is a mismatch against the $-$ sequence only,
the symbols $\downarrow_{/i}^+$ signify that it is a mismatch against the
$+$ only, and the symbol $\Downarrow$ signifies that it is a mismatch
against both. As before, every other nucleotide is replaced by the symbol
$\square$, and the terminator $|$ is appended to the end of the read. We
again define reads as sequences of segments, except that now the
terminators are the symbols $\downarrow_{/j}^-$, $\downarrow_{/j}^+$ and
$\Downarrow$. The tail, as usual, is terminated by the symbol $|$.

The number $i$ in the symbol $\downarrow_{/i}^-$ indicates the match
length of the $+$ sequence. Likewise, the number $i$ in the symbol
$\downarrow_{/i}^+$ indicates the match length of the $-$ sequence. For
instance, the symbol $\downarrow_{/7}^-$ indicates that the nucleotide is a
mismatch against the $-$ sequence, that it is a match for the $+$
sequence, and that the six previous nucleotides were also a match for the
$+$ sequence. The terminators thus encode the local state of the read.

Figure~\ref{fig:dual} shows an example of read in the dual encoding. The $+$
and $-$ sequences are represented below the read, with matches represented
as open squares and mismatches as closed squares. It is visible from this
example that symbols $\downarrow^-_{/i}$ and $\downarrow^+_{/i}$ alternate
whenever the mismatches hit different sequences . The symbol $\Downarrow$
occurs only when a nucleotide is a double mismatch.

@@

We assume that for each nucleotide, $a$ is the probability that the read
matches both sequences, $b$ is the probability that it matches only the
$+$ sequence, $c$ is the probability that it matches only the $-$ sequence
and $d$ is the probability that it matches none. Obviously $a+b+c+d=1$.
These definitions allow us to specify the weighted generating function of
the symbols and of the segments. To specify the transfer matrix, we only
have to make sure that we eliminate the combinations that create a match
of size $\gamma$ or more for any of the two sequences. For notational
convenience, we define

@@

r_i^+(z) = cz \cdot (az)^i, \\
r_i^-(z) = bz \cdot (az)^i, \\
R_i(z) = dz \cdot \big( 1 + az + \ldots + (az)^i \big), \\

@@

The term of interest is the top right entry of
$\tilde{M}_0(z)\cdot(I-\tilde{M}_0(z))^{-1}$. To see why, observe that
every read can be prepended by $\Downarrow$ segments and only by those
(every other terminator would imply that one of the two sequences has a
nonzero match size at the start of the read). Thus reads are precisely the
sequences of segments that can follow the symbol $\Downarrow$ and that are
terminated by a tail, whose weighted generating function is the top right
entry of the matrix.

$\tilde{M}_0(z)$ is too complex to allow a closed expression of
$\tilde{M}_0(z)\cdot(I-\tilde{M}_0(z))^{-1}$ to be computed. It is
actually easier to proceed as in section~\ref{sec:example_skip} and to
compute the powers of $\tilde{M}_0(z)$ (using the arithmetic of truncated
polynomials) up to a given threshold value. Since each segment except the
tail contains a mismatch against at least one sequence, we define
$\tilde{p}$ as the upper bound on the probability of a mismatch $\tilde{p}
= \max\{b,c,d\}$. The updated formula (\ref{eq:bound}) now gives an upper
bound of the probability that a read of size $k$ contains $s$ or more
mismatches as
\begin{equation*}
Pr(X \geq s) \leq \exp \left( (s-k)\log \frac{k-s}{k(1-\tilde{p})} -s\log
\frac{s}{k\tilde{p}} \right).
\end{equation*}

@@

With these values, we can fully specify the matrix $\tilde{M}_0(z)$, and
compute its powers until the upper bound is lower than a set fraction
$\varepsilon$ of the current value of $a_k$, which gives an accurate
estimate of $P(\overline{S}_0 \cap \overline{S}_1)$.

\begin{remark}
Observe that expression (\ref{eq:Pnull}) is the probability of null
seeding. The method described above thus allows us to compute this
probability at no additional cost. This probablity is less relevant than
the probabilities that the seeding process is on target or off target, but
at times, it may be useful to know the probability that a read is
not mappable, especially when reads are relatively short.
\end{remark}

@@

To compute the probability that the seeding process is off target when
using skip seeds, we observe that the logic of section~\ref{sec:offexact}
can be transposed with few modifications. In particular, the probability
can be computed through expression (\ref{eq:Poff}), where $S_0$ is the
event that the read has a \emph{skip} seed for the target and $S_1$ that
it has a \emph{skip} seed for the first duplicate.

@@

Once again, we start by recoding the reads in a specialized alphabet. We
define the skip-$n$ dual alphabet $\tilde{\mathcal{A}}_n = \{\square, |,
\Downarrow_0, \Downarrow_1, \Downarrow_2, \ldots, \Downarrow_n,
\downarrow^-_{/1}, \downarrow^-_{/2}, \ldots, \downarrow^+_{/1},
\downarrow^+_{/2}, \ldots\}$. The symbols $\square$, $|$,
$\downarrow^-_{/i}$ and $\downarrow^+_{/i}$ have the same meaning as in
the dual alphabet of section~\ref{sec:dual}. The symbols $\Downarrow_i$
indicate that both sequences have match length 0 and that the next
non-skipped position is $i$ nucleotides downstream.

@@

\section{Random seeds}
\label{sec:random_seeds}

\section{True versus best location}
\label{sec:truevsbest}

@@

@@

\subsection{Derivation of something important}
\label{app_skip}

Easy.

observe that the terms $M_n(z)^{k+2}, M_n(z)^{k+3},
\ldots$ have no influence on the coefficients of the weighted generating
function $a_0, a_1, \ldots, a_k$.

Indeed, a read of size $k$ has at most $k+1$ segments (including the
obligatory tail, which can have size 0). Since $M(z)^{s+1}$ contains the
weighted generating functions of reads with exactly $s+1$ segments
including the tail, $a_k$ cannot depend on $M(z)^{k+2}, M(z)^{k+3},
\ldots$ More formally, we can prove by induction that all the entries of
$M(z)^k$ are divisible by $z^{k-1}$, showing that the contribution of
$M(z)^{k+2} + M(z)^{k+3} + \ldots$ to $a_0 + a_1z + \ldots +a_kz^k$ is
strictly $0$.

>>>>

We need to know whether a nucleotide is a sequencing error, but this time
we also need to know its phase in the repeated cycles of skipped
positions. For this, we define the skip-$n$ alphabet $\mathcal{A}_n =
\{\square, |, \Downarrow_0, \Downarrow_1, \ldots, \Downarrow_n \}$. Again,
the symbol $\square$ represents a correct nucleotide and the symbol $|$ is
a terminator added at the end of the read. The symbols $\Downarrow_j$ ($ 0
\leq j \leq n$) represent sequencing errors and $j$ indicates the number
of nucleotides till the next non-skipped position (\textit{i.e.} $j=0$ for
nucleotides immediately before a non-skipped position and $j=n$ for
nucleotides at a non-skipped position).

@@

Finally, the expression of $M_n(z)$ the transfer matrix of reads in the
skip-$n$ alphabet is
\begin{equation}
\label{eq_matrix_Mn}
\begin{blockarray}{cccccc}
   & \Dn{0} & \Dn{1} & \ldots & \Dn{n} & \nd \\
\begin{block}{c[ccccc]}
\Dn{0} & H_{0,0}(z) & H_{0,1}(z) & \ldots & H_{0,n}(z) & J_0(z) \\
\Dn{1} & H_{1,0}(z) & H_{1,1}(z) & \ldots & H_{1,n}(z) & J_1(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\Dn{n} & H_{n,0}(z) & H_{n,1}(z) & \ldots & H_{n,n}(z) & J_n(z) \\
\nd & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray}.
\end{equation}

@@

By construction, the Taylor expansion of the top right term in the matrix
$M_n(z)\cdot(I-M_n(z))^{-1}$ contains the the probabilities of interest.
More specifically, if the Taylor series of this term is $a_0 + a_1 z + a_2
z^2 + \ldots$, then $a_k$ is the probability that a read of size $k$
contains no skip-$n$ seed of minimum size $\gamma$.

But $M_n(z)$ is too complex to find a closed expression for $M_n(z) \cdot
(I-M_n(z))^{-1}$ or its top right term. Instead, we return to the
definition $M_n(z) + M_n(z)^2 + \ldots$ and show in
appendix~\ref{app_skip} that the coefficient of interest, $a_k$, only
depends on the first $k+1$ terms of the sum. So for reads of size $k$ or
lower, we only need to compute the matrix $M_n(z) + M_n(z)^2 + \ldots +
M_n(z)^{k+1}$ and work out the Taylor expansion of the top right term.

@@

Using the formula above, we can thus bound the probability that a read has
$s+1$ or more segments. We compute $M_n(z) + M_n(z)^2 + \ldots
+M_n(z)^{s+1}$ where the weighted generating functions have been replaced
by truncated polynomials and we extract the top right entry. When the
right-hand side of expression (\ref{eq:bound}) is lower than a set
fraction $\varepsilon$ of the current value of $a_k$, we stop the
computations. Typically $\varepsilon = 0.01$ so this method ensures that
the probabilities that a read of size $k$ has no on-target skip seed are
accurate to within 1\%.

@@

We now turn our attention to the problem of computing the probability that
the seeding process is off target when using exact seeds---recall from
section~\ref{sec:twotypes} that off-target seeding means that the
candidate set contains a duplicate but not the target.

If there is no duplicate (\textit{i.e.} $N = 0$), seeding cannot be off
target, it can only be on target or null. So from here we assume that the
target has $N \geq 1$ duplicates. Let $S_0$ denote the event that there is
an on-target seed and let $S_j$ denote the event that there is a seed for
the $j$-th duplicate. We are thus interested in computing
$P\big(\overline{S}_0 \cap (S_1 \cup \ldots \cup S_N)\big)$, where
$\overline{S}_j$ denotes the complement of the event $S_j$. First observe
that
\begin{equation}
\label{eq:pcomp}
P\big(\overline{S}_0 \cap (S_1 \cup \ldots \cup S_N)\big) =
P(\overline{S}_0) - P\big(\overline{S}_0 \cap \overline{S}_1 \cap \ldots
\cap \overline{S}_N\big).
\end{equation}

Since the duplicates are assumed to evolve independently of each other and
through the same mutagenesis process, the events $S_j$ ($1 \leq j \leq N$)
are independent and identically distributed conditionally on
$\overline{S}_0$. We can thus write
\begin{eqnarray}
\notag
P(\overline{S}_0 \cap \ldots \cap \overline{S}_N) &=&
  P(\overline{S}_0) \cdot P(\overline{S}_1 \cap \ldots \cap
  \overline{S}_N | \overline{S}_0) \\
\label{eq:Pnull}
  &=& P(\overline{S}_0) \cdot P(\overline{S}_1 | \overline{S}_0)^N =
  P(\overline{S}_0) \cdot \left( \frac{P(\overline{S}_0 \cap
\overline{S}_1)}{P(\overline{S}_0)} \right)^N.
\end{eqnarray}
Combinging the two equations above, we obtain
\begin{equation}
\label{eq:Poff}
P\big(\overline{S}_0 \cap (S_1 \cup \ldots \cup S_N)\big) =
P(\overline{S}_0) - P(\overline{S}_0) \cdot \left( \frac{P(\overline{S}_0
\cap \overline{S}_1)}{P(\overline{S}_0)} \right)^N.
\end{equation}

Hence, the probability that seeding is off target is a function of just
two quantitites: $P(\overline{S}_0)$ and $P(\overline{S}_0 \cap
\overline{S}_1)$. The first is the probability that the read has no seed
for the target, which we have already computed in
section~\ref{sec:example_exact} using recursive expression
(\ref{eq:recur}). We now need to find a way to compute $P(\overline{S}_0
\cap \overline{S}_1)$.

\begin{remark}
Observe that expression (\ref{eq:Pnull}) is the probability of null
seeding (the read contains no seed for the target or any of its
duplicates). Since it is also a function of just $P(\overline{S}_0)$ and
$P(\overline{S}_0 \cap \overline{S}_1)$, it can be computed at no
additional cost. This probablity is less relevant than the probabilities
that the seeding process is on target or off target, but at times, it may
be useful to know the probability that a read is not mappable, especially
when reads are relatively short.
\end{remark}

@@

$P(\overline{S}_0 \cap \overline{S}_1)$ is the probability that the read
has no seed for the target or for the first duplicate---numbering is
arbitrary here, the first duplicate can be any fixed duplicate. As
illustrated in section~\ref{sec:symbolic}, we first recode the reads using
a specialized alphabet to simplify the problem.

It will be useful to consider a more general problem where we
have two sequences of interest labelled $+$ and $-$. The $+$ sequence
stands for the target and that the $-$ sequence stands for its duplicate.
We then define the dual alphabet $\tilde{\mathcal{A}}_0 = \{\square, |,
\downarrow_{/1}^-, \downarrow_{/2}^-, \ldots, \downarrow_{/1}^+,
\downarrow_{/2}^+, \ldots, \Downarrow\}$. The symbols $\downarrow_{/i}^-$
($i \geq 1$) signify that the nucleotide is a mismatch against the $-$
sequence only, the symbols $\downarrow_{/i}^+$ ($i \geq 1$) signify that
it is a mismatch against the $+$ sequence only, and the symbol
$\Downarrow$ signifies that it is a mismatch against both. As before,
every other nucleotide is replaced by the symbol $\square$, and the
terminator $|$ is appended to the end of the read. We again define reads
as sequences of segments, except that now the terminators are the symbols
$\downarrow_{/i}^-$, $\downarrow_{/i}^+$ and $\Downarrow$ ($i \geq 1$).
The tail, as usual, is terminated by the symbol $|$.

The index $i$ in the symbol $\downarrow_{/i}^-$ indicates the match
length of the $+$ sequence. Likewise, the index $i$ in the symbol
$\downarrow_{/i}^+$ indicates the match length of the $-$ sequence. For
instance, the symbol $\downarrow_{/7}^-$ indicates that the nucleotide is a
mismatch against the $-$ sequence, that it is a match for the $+$
sequence, and that the six preceding nucleotides were also a match for the
$+$ sequence (but the nucleotide before that was a mismatch against the
$+$ sequence). The terminators thus encode the local state of the read.

Figure~\ref{fig:dual} shows an example of read in the dual encoding. The
$+$ and $-$ sequences are shown below the read, with matches represented
as open squares and mismatches as closed squares. It is visible from this
example that symbols $\downarrow^-_{/i}$ and $\downarrow^+_{/i}$ alternate
whenever the mismatches hit different sequences. The symbol $\Downarrow$
occurs only when a nucleotide is a double mismatch.

@@

Let us assume that for each nucleotide, $a$ is the probability that the
read matches both sequences, $b$ is the probability that it matches only
the $+$ sequence, $c$ is the probability that it matches only the $-$
sequence and $d$ is the probability that it matches neither. Obviously
$a+b+c+d=1$. These definitions allow us to specify the weighted generating
function of the symbols and of the segments. To specify the transfer
matrix, we only have to make sure that we eliminate the combinations that
create a match of size $\gamma$ or more for any of the two sequences. For
notational convenience, we define

@@

r_i^+(z) = (az)^i cz, \\
r_i^-(z) = (az)^i bz, \\
R_i(z) = \big( 1 + az + \ldots + (az)^i \big) dz, \\

@@

As before, the term of interest is the top right entry of
$\tilde{M}_0(z)\cdot(I-\tilde{M}_0(z))^{-1}$. To see why, observe that
every read can be prepended by $\Downarrow$ segments and only by those
(every other terminator would imply that one of the two sequences has a
nonzero match size at the start of the read). Thus reads are precisely the
sequences of segments that can follow the symbol $\Downarrow$ and that are
terminated by a tail, the weighted generating function of which is located
in the top right entry of the matrix.

$\tilde{M}_0(z)$ is too complex to compute a closed expression of
$\tilde{M}_0(z)\cdot(I-\tilde{M}_0(z))^{-1}$. It is actually easier to
proceed as in section~\ref{sec:example_skip} and to compute the powers of
$\tilde{M}_0(z)$ up to a finite value. This is done once again using the
arithmetic of truncated polynomials. Since each segment except the tail
contains a mismatch against at least one sequence, we define $\tilde{p}$
as the upper bound on the probability of a mismatch, \textit{i.e.}
$\tilde{p} = \max\{b,c,d\}$. The updated formula (\ref{eq:bound}) now
gives an upper bound of the probability that a read of size $k$ contains
$s$ or more mismatches as
\begin{equation*}
Pr(X \geq s) \leq \exp \left( (s-k)\log \frac{k-s}{k(1-\tilde{p})} -s\log
\frac{s}{k\tilde{p}} \right).
\end{equation*}

@@

With these values, we can fully specify the matrix $\tilde{M}_0(z)$, and
compute its powers until the right-hand side of the expression above is
lower than a set fraction $\varepsilon$ of the current value of $a_k$,
which gives an accurate estimate of $P(\overline{S}_0 \cap
\overline{S}_1)$.

@@

To compute the probability that the seeding process is off target when
using skip seeds, we observe that the logic of section~\ref{sec:offexact}
can be transposed with few modifications. In particular, the probability
can be computed through expression (\ref{eq:Poff}), where $S_0$ is the
event that the read has a \emph{skip} seed for the target (instead of an
exact seed) and $S_1$ is the event that it has a \emph{skip} seed for the
first duplicate (instead of an exact seed).

@@

Once again, we start by recoding the reads in a specialized alphabet. We
define the skip-$n$ dual alphabet $\tilde{\mathcal{A}}_n = \{\square, |,
\Downarrow_0, \Downarrow_1, \Downarrow_2, \ldots, \Downarrow_n,
\downarrow^-_{/1}, \downarrow^-_{/2}, \ldots, \downarrow^+_{/1},
\downarrow^+_{/2}, \ldots\}$. The symbols $\square$, $|$,
$\downarrow^-_{/i}$ and $\downarrow^+_{/i}$ ($i \geq 1$) have the same
meaning as in the dual alphabet of section~\ref{sec:dual}. The symbols
$\Downarrow_i$ ($0 \leq i \leq n$) indicate that both sequences have match
length 0 and that the next non-skipped position is $i$ nucleotides
downstream.

@@

\section{Practical considerations}

\subsection{True versus best location}
\label{sec:truevsbest}

@@

\subsection{Non-specific seeds}
\label{sec:random_seeds}

\subsection{Realistic error patterns}

This section will come later.

@@

\subsection{Proof that $a_k$ is independent of $M_n(z)^m$ for $m > k+1$}
\label{app_skip}

In section~\ref{sec:example_skip} we considered the transfer matrix
$M_n(z)$ of reads containing no skip-$n$ seed. The quantity of interest
$a_k$ is the coefficent of $z^k$ in the Taylor expansion of the top right
term of $M_n(z) + M_n(z)^2 + M_n(z)^3 + \ldots$ We claimed that the
coefficients $a_0, a_1, \ldots, a_k$ are independent of $M_n(z)^{k+2},
M_n(z)^{k+3}, \ldots$ so that they can be computed from the finite sum
$M_n(z) + M_n(z)^2 + \ldots + M_n(z)^{k+1}$.

We proceed by induction and show that for $k \geq 1$ all the entries of
$M_n(z)^k$ are mutliples of $z^{k-1}$. For $k = 1$ the statement is
trivial and there is nothing to show. Now assume that for some $k > 1$,
all the entries of $M_n(z)^k$ are multiples of $z^{k-1}$. Observe that
from the definition of $M_n(z)$ in expression (\ref{eq_matrix_Mn}), the
last row of $M_n(z)^m$ consists of zeros only for all $m \geq 1$. So in
the matrix product $M_n(z) \cdot M_n(z)^k$, the polynomials $J_0(z),
\ldots, J_n(z)$ in the last column of $M_n(z)$ are all multiplied by 0.
The remaining terms of $M_n(z)$ are multiples of $z$, so applying the
induction hypothesis shows that all the terms of $M_n(z)^{k+1}$ are
multiples of $z^k$.

Since all the terms of $M(z)^m$ are multiples of $z^{k+1}$ for $m \geq
k+1$, the coefficient of $z^k$ is 0 in the top right term of every matrix
$M(z)^m$ for $m \geq k+1$. As a result, the coefficient of $z^k$ in the
top right entry of $M_n(z) + M_n(z)^2 + M_n(z)^3 + \ldots$ is the same as
in $M_n(z) + M_n(z)^2 + \ldots + M_n(z)^{k+1}$, as we had to demonstrate.
The same rationale holds for the coefficients of $z^m$ $(0 \leq m < k)$,
so finally the coefficients $a_0, a_1, \ldots, a_k$ can all be computed
from the top right entry of $M_n(z) + M_n(z)^2 + \ldots + M_n(z)^{k+1}$ as
claimed.
